AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide

### 2021-08-30 publications/dora_outlier_detection_iaai.tex

\begin{abstract}
AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors. To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions.
\end{abstract}

\section{Introduction}
\begin{itemize}
\item Framing of outlier detection as a domain-agnostic problem that is well-suited to cross-disciplinary applications
\item Example applications of outlier detection across science application domains
\item Challenges/barriers to entry for domain scientists who could benefit from outlier detection 
\item DORA as a solution to lower these barriers to entry and help integrate outlier detection methods into existing scientific workflows across variety of domains
\end{itemize}

@@

\begin{itemize}
\item DORA configurable pipeline approach 
\item data types
\item feature extraction
\item ranking algorithms
\item results modules
\end{itemize}

@@

\subsection{Datasets}

Four use case datasets; additional benchmark datasets (e.g. FashionMNIST/MNIST)?

\subsection{Experiment setup}

Experiments for each of the use cases; evaluation criteria

\section{Results}

Experiment results

\section{Discussion}
\begin{itemize}
\item Discuss results from experiments
\item Challenges of evaluating results in outlier detection?
\item Downstream use of results in science applications
\end{itemize}

\section{Conclusions}
Summarize conclusions and plans for future work

Path to deployment

@@

\section{ Acknowledgments}
JPL CIF, SMD etc.

>>>>

\begin{abstract}
Automatic detection of out-of-distribution (OOD) samples or features is
universally needed when working with scientific datasets. For example,
OOD detection methods can be used for cleaning datasets or flagging
novel samples in order to guide instrument acquisition or scientific analysis.
We present Domain-agnostic Outlier Ranking Algorithms (DORA) to provide
a configurable pipeline that makes it easy for scientists to apply and evaluate
outlier detection methods in a variety of domains. DORA allows users to 
configure outlier detection experiments by specifying the location of their 
dataset(s), the input data type, feature extraction methods, and which 
algorithms should be applied. DORA supports image, raster, time series, 
or feature vector input data types and outlier detection methods that include
 Isolation Forest, Discovery via Eigenbasis Modeling of Uninteresting Data 
 (DEMUD), principal component analysis (PCA), Reed-Xiaoli (RX) detector, 
 Local RX, negative sampling, and probabilistic autoencoder. Given a set of
  data samples, each algorithm assigns an outlier score to each sample. DORA
   provides results organization and visualization modules to help users 
   process the results, including sorting samples by outlier score, evaluating 
   outlier recall for a set of known/labeled outliers, clustering groups of 
   similar outliers together, and causal graphs. To demonstrate the 
   cross-domain applicability of DORA, we conducted outlier detection
    experiments for four use cases: 1) cleaning noisy ground-truth datasets
     (Earth Science), 2) detecting volcanic thermal anomalies (Earth Science), 
     3) flagging non-astrophysical objects in the Dark Energy Survey 
     (astrophysics), and 4) novelty-guided onboard targeting of instruments
      for Mars rovers (planetary science). [We showed that ...]
\end{abstract}

\section{Introduction}
\begin{itemize}
\item Framing of outlier detection as a domain-agnostic problem that is well-
suited to cross-disciplinary applications
\item Example applications of outlier detection across science application 
domains
\item Challenges/barriers to entry for domain scientists who could benefit from
 outlier detection 
\item DORA as a solution to lower these barriers to entry and help integrate 
outlier detection methods into existing scientific workflows across variety of 
domains
\end{itemize}

@@

\subsection{DORA configurable pipeline} Steven please add details about the 
approach/design choices for this pipeline. Mention that additional data types, 
feature extraction methods, ranking algorithms, and results organization and 
visualization modules can easily be added and integrated if they follow the API. 
Hannah to add architecture figure. Add descriptions of each layer of modules: 
\paragraph{Data types} Hannah
\paragraph{Feature extraction} Hannah 
\paragraph{Ranking algorithms} We implemented 7 algorithms for scoring and 
ranking samples by outlierness. [Describe each algorithm briefly and how they 
complement each other.]
\paragraph{Results organization} Eric, Umaa, Steven, Vinay
\paragraph{Outlier visualization} Jake?

@@

\subsection{Datasets}
Describe the datasets overall and our motivation for choosing these. Maybe add
 a table with number of samples, number of reference known outliers (if 
 available), etc, if room?

\subsubsection{Astrophysical objects in Dark Energy Survey}
Umaa

\subsubsection{Targets in Mars rover images}
Kiri

\subsubsection{Satellite time series for ground-truth observations}
Hannah: Describe the Earth crop field data set and applications

\subsubsection{Volcanic thermal anomalies}

\subsubsection{MNIST and FashionMNIST}
We added this to have a traditional benchmark and use case demonstration.

\subsection{Results}
Discuss the overall approach to the setup of the experiments. In each 
subsection discuss specific setup details (including evaluation metrics) for 
each dataset then the results from the experiments.

\subsubsection{Astrophysical objects in Dark Energy Survey}
Umaa

\subsubsection{Targets in Mars rover images}
Kiri

\subsubsection{Satellite time series for ground-truth observations}
Hannah: Describe the Earth crop field data set and applications

\subsubsection{Volcanic thermal anomalies}

\subsubsection{MNIST and FashionMNIST}
We added this to have a traditional benchmark and use case demonstration.

\section{Discussion}
\begin{itemize}
\item Discuss results from experiments
\item Challenges of evaluating results in outlier detection?
\item Downstream use of results in science applications/Path to deployment
\end{itemize}

\section{Conclusions}
Summarize conclusions and plans for future work, including public 
pip-installable package

[Note: these datasets could also be used by other researchers as OOD
 benchmark tasks?]

@@

\section{ Acknowledgments}
Funding/support: JPL CIF, SMD etc. Acknowledge people who contributed to 
datasets e.g. FAO, Matt Pritchard, DES, other CIF team members?
